The role of emotional factors in face processing abilities in Autism Spectrum Conditions

### 2023-07-09 manuscript.tex
Finished analysis and added autobuild

We will planned to include some simple correlational analyses, descriptive statistics, and t-tests to further clarify any group differences, and directions of effect. We also examined the group differences of participants who meet the commonly used threshold for alexithymia (a score of 61 of above on the TAS-20), and those who do not, as previous research indicates we will also find a difference between these groups (\protect\hyperlink{ref-cook2013a}{Cook et al., 2013}). This was done via a series of non-parametric tests, to explore whether those who are above the threshold are associated with higher levels of depression, anxiety, autistic traits, and poor facial emotion and identity recognition.

@@

In our first series of regressions, a continuous measure of autistic traits, the AQ28, was entered in the final step. Table \ref{tab:table3} summarises the results of this regression series. For emotion attribution thresholds, identity attribution thresholds, and identity PSE, no model was significant at any step. For emotion PSE, the regression model was significant at steps 1s and 2 (although the addition of the TAS in step 2 did not lead to a significant increase in the variance explained). The only significant predictor was gender, specifically the dummy variable indicating ``other'' gender. See sub-section ``The role of gender'' for follow up tests on gender effects on emotion PSE. Table 4 summarises the coefficients for the regressions with significant models.

@@

Table 5 summarises the results of the regression series that considered the role of autistic status. For emotion attribution thresholds, identity attribution thresholds, and identity PSE, no model was significant at any step. For emotion PSE, the regression model was significant at steps 1, 2 and 4, although beyond model 1 no additional steps added significant explained variance. Table 6 summarises the coefficients for the regressions with significant models.

Table 7 summarises the results of the regression series that considered the role of self-identification. Similar to the regression series reported above, emotion attribution thresholds, identity attribution thresholds, and identity PSE, no model was significant at any step. For emotion PSE, the regression model was significant at steps 1, 2 and 4, although beyond model 1 no additional steps added significant explained variance. Table 8 summarises the coefficients for the regressions with significant models.

@@

We compared participants who scored above the standard threshold for ``high'' alexithymic traits to those who scored below this threshold on our face processing measures, autistic traits, and anxiety and depression scores. We again removed outliers: this was done using the thresholds used for regression series 1 (see Supplementary Materials for results when all participants including outliers are included). Given that the face processing variables, AQ scores and HADS scores were not normally distributed, these comparisons were conducted using the non-parametric Mann-Whitney U tests. Table 9 includes the statistics for these comparisons.

Comparing participants who scored above and below the high alexithymia cut-offs on the face processing variables, no comparisons reached significance (see Table 9). Highly alexithymic participants had higher autistic traits, and lower anxiety scores. Depression scores were not significantly different between the group.

>>>>

We planned to include some simple correlational analyses, descriptive statistics, and t-tests to further clarify any group differences, and directions of effect. We also examined the group differences of participants who meet the commonly used threshold for alexithymia (a score of 61 of above on the TAS-20), and those who do not, as previous research indicates we will also find a difference between these groups (\protect\hyperlink{ref-cook2013a}{Cook et al., 2013}). This was done via a series of non-parametric tests, to explore whether those who are above the threshold are associated with higher levels of depression, anxiety, autistic traits, and poor facial emotion and identity recognition.

@@

In our first series of regressions, a continuous measure of autistic traits, the AQ28, was entered in the final step. Table \ref{tab:table3} summarises the results of this regression series. For emotion attribution thresholds, identity attribution thresholds, and identity PSE, no model was significant at any step. For emotion PSE, the regression model was significant at steps 1s and 2 (although the addition of the TAS in step 2 did not lead to a significant increase in the variance explained). The only significant predictor was gender, specifically the dummy variable indicating ``other'' gender. See sub-section ``The role of gender'' for follow up tests on gender effects on emotion PSE. Table \ref{tab:table4} summarises the coefficients for the regressions with significant models.

@@

Table \ref{tab:table5} summarises the results of the regression series that considered the role of autistic status. For emotion attribution thresholds, identity attribution thresholds, and identity PSE, no model was significant at any step. For emotion PSE, the regression model was significant at steps 1, 2 and 4, although beyond model 1 no additional steps added significant explained variance. Table \ref{tab:table6} summarises the coefficients for the regressions with significant models.

\caption{\label{tab:table5}Regression Series 2 (Autistic vs Non-autistic Group status entered in Step 4).}
\centering
\fontsize{5}{7}\selectfont
\begin{tabular}[t]{lllllllll}
\toprule
\multicolumn{1}{l}{} & \multicolumn{2}{l}{Emotion Attribution} & \multicolumn{2}{l}{Identity Attribution} & \multicolumn{2}{l}{Emotion PSE} & \multicolumn{2}{l}{Identity PSE} \\
\multicolumn{1}{l}{} & \multicolumn{2}{l}{Total N (ASC,TD) = 174 (77, 97)} & \multicolumn{2}{l}{Total N (ASC,TD) = 170 (75, 95)} & \multicolumn{2}{l}{Total N (ASC,TD) = 174 (77, 97)} & \multicolumn{2}{l}{Total N (ASC,TD) = 176 (78, 98)} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
 &  & F change, &  & F change, &  & F change, &  & F change,\\
 & F (Df), p-value & p-value & F (Df), p-value & p-value & F (Df), p-value & p-value & F (Df), p-value & p-value\\
\midrule
Model 1 & 1.63 (5,168), .156 & 1.63, .156 & 0.19 (5,164), .968 & 0.19, .968 & 2.4 (5,168), .039 & 2.4, .039 & 1.46 (5,170), .207 & 1.46, .207\\
Model 2 & 1.98 (6,167), .071 & 3.63, .059 & 0.57 (6,163), .753 & 2.49, .116 & 2.34 (6,167), .034 & 1.96, .163 & 1.69 (6,169), .126 & 2.79, .097\\
Model 3 & 1.57 (8,165), .136 & 0.39, .676 & 0.93 (8,161), .49 & 2, .139 & 1.96 (8,165), .055 & 0.83, .439 & 1.26 (8,167), .27 & 0.01, .988\\
\addlinespace
Model 4 & 1.59 (9,164), .122 & 1.66, .199 & 1.07 (9,160), .386 & 2.13, .147 & 2.11 (9,164), .031 & 3.15, .0779 & 1.14 (9,166), .337 & 0.26, .609\\
\bottomrule
\end{tabular}
\end{table}

\caption{\label{tab:table6}Coefficients for Emotion PSE Regression series 2.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Model &  & Beta & t & p-value\\
\midrule
1 & (Constant) &  & 39.929 & <.001\\
 & Age & -0.119 & -1.588 & .114\\
 & Male gender & -0.134 & -1.717 & .088\\
 & Other gender & -0.218 & -2.792 & .006\\
 & Asian ethnicity & 0.053 & 0.712 & .477\\
\addlinespace
 & Mixed ethnicity & 0.018 & 0.239 & .812\\
2 & (Constant) &  & 23.210 & <.001\\
 & Age & -0.136 & -1.796 & .074\\
 & Male gender & -0.141 & -1.803 & .073\\
 & Other gender & -0.225 & -2.887 & .004\\
\addlinespace
 & Asian ethnicity & 0.061 & 0.809 & .419\\
 & Mixed ethnicity & 0.01 & 0.130 & .896\\
 & TAS-TOTAL & 0.106 & 1.402 & .163\\
3 & (Constant) &  & 13.604 & <.001\\
 & Age & -0.138 & -1.808 & .072\\
\addlinespace
 & Male gender & -0.15 & -1.861 & .065\\
 & Other gender & -0.225 & -2.880 & .005\\
 & Asian ethnicity & 0.049 & 0.652 & .515\\
 & Mixed ethnicity & 0.001 & 0.008 & .993\\
 & TAS-TOTAL & 0.108 & 1.347 & .18\\
\addlinespace
 & HADS Anxiety & 0.026 & 0.322 & .748\\
 & HADS Depression & 0.093 & 1.230 & .22\\
4 & (Constant) &  & 13.757 & <.001\\
 & Age & -0.229 & -2.502 & .013\\
 & Male gender & -0.109 & -1.297 & .196\\
\addlinespace
 & Other gender & -0.247 & -3.141 & .002\\
 & Asian ethnicity & 0.051 & 0.675 & .5\\
 & Mixed ethnicity & -0.005 & -0.063 & .95\\
 & TAS-TOTAL & 0.066 & 0.792 & .429\\
 & HADS Anxiety & 0.048 & 0.585 & .559\\
\addlinespace
 & HADS Depression & 0.075 & 0.987 & .325\\
 & Autism Group (autistic or non-autistic) & 0.183 & 1.774 & .078\\
\bottomrule
\end{tabular}
\end{table}

Table \ref{tab:table7} summarises the results of the regression series that considered the role of self-identification. Similar to the regression series reported above, emotion attribution thresholds, identity attribution thresholds, and identity PSE, no model was significant at any step. For emotion PSE, the regression model was significant at steps 1, 2 and 4, although beyond model 1 no additional steps added significant explained variance. Table \ref{tab:table8} summarises the coefficients for the regressions with significant models.

\caption{\label{tab:table7}Regression series 3 (Autistic vs Non-autistic vs Self-identifying status entered in Step 4).}
\centering
\fontsize{5}{7}\selectfont
\begin{tabular}[t]{lllllllll}
\toprule
\multicolumn{1}{l}{} & \multicolumn{2}{l}{Emotion Attribution} & \multicolumn{2}{l}{Identity Attribution} & \multicolumn{2}{l}{Emotion PSE} & \multicolumn{2}{l}{Identity PSE} \\
\multicolumn{1}{l}{} & \multicolumn{2}{l}{Total N (ASC,TD) = 174 (77, 97)} & \multicolumn{2}{l}{Total N (ASC,TD) = 170 (75, 95)} & \multicolumn{2}{l}{Total N (ASC,TD) = 174 (77, 97)} & \multicolumn{2}{l}{Total N (ASC,TD) = 176 (78, 98)} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
 &  & F change, &  & F change, &  & F change, &  & F change,\\
 & F (Df), p-value & p-value & F (Df), p-value & p-value & F (Df), p-value & p-value & F (Df), p-value & p-value\\
\midrule
Model 1 & 1.63 (5,168), .156 & 1.63, .156 & 0.19 (5,164), .968 & 0.19, .968 & 2.4 (5,168), .039 & 2.4, .039 & 1.46 (5,170), .207 & 1.46, .207\\
Model 2 & 1.98 (6,167), .071 & 3.63, .059 & 0.57 (6,163), .753 & 2.49, .116 & 2.34 (6,167), .034 & 1.96, .163 & 1.69 (6,169), .126 & 2.79, .097\\
Model 3 & 1.57 (8,165), .136 & 0.39, .676 & 0.93 (8,161), .49 & 2, .139 & 1.96 (8,165), .055 & 0.83, .439 & 1.26 (8,167), .27 & 0.01, .988\\
\addlinespace
Model 4 & 1.5 (10,163), .145 & 1.17, .312 & 0.98 (10,159), .461 & 1.17, .314 & 1.91 (10,163), .047 & 1.66, .193 & 1.02 (10,165), .427 & 0.14, .871\\
\bottomrule
\end{tabular}
\end{table}

\caption{\label{tab:table8}Coefficients for Emotion PSE Regression Series 3.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Model &  & Beta & t & p-value\\
\midrule
1 & (Constant) &  & 39.929 & <.001\\
 & Age & -0.119 & -1.588 & .114\\
 & Male gender & -0.134 & -1.717 & .088\\
 & Other gender & -0.218 & -2.792 & .006\\
 & Asian ethnicity & 0.053 & 0.712 & .477\\
\addlinespace
 & Mixed ethnicity & 0.018 & 0.239 & .812\\
2 & (Constant) &  & 23.210 & <.001\\
 & Age & -0.136 & -1.796 & .074\\
 & Male gender & -0.141 & -1.803 & .073\\
 & Other gender & -0.225 & -2.887 & .004\\
\addlinespace
 & Asian ethnicity & 0.061 & 0.809 & .419\\
 & Mixed ethnicity & 0.01 & 0.130 & .896\\
 & TAS-TOTAL & 0.106 & 1.402 & .163\\
3 & (Constant) &  & 13.604 & <.001\\
 & Age & -0.138 & -1.808 & .072\\
\addlinespace
 & Male gender & -0.15 & -1.861 & .065\\
 & Other gender & -0.225 & -2.880 & .005\\
 & Asian ethnicity & 0.049 & 0.652 & .515\\
 & Mixed ethnicity & 0.001 & 0.008 & .993\\
 & TAS-TOTAL & 0.108 & 1.347 & .18\\
\addlinespace
 & HADS Anxiety & 0.026 & 0.322 & .748\\
 & HADS Depression & 0.093 & 1.230 & .22\\
4 & (Constant) &  & 13.719 & <.001\\
 & Age & -0.237 & -2.532 & .012\\
 & Male gender & -0.108 & -1.285 & .201\\
\addlinespace
 & Other gender & -0.25 & -3.160 & .002\\
 & Asian ethnicity & 0.048 & 0.638 & .524\\
 & Mixed ethnicity & -0.003 & -0.039 & .969\\
 & TAS-TOTAL & 0.065 & 0.780 & .436\\
 & HADS Anxiety & 0.048 & 0.580 & .563\\
\addlinespace
 & HADS Depression & 0.076 & 0.996 & .321\\
 & Formal Diagnosis (vs non-autistic) & 0.173 & 1.710 & .089\\
 & Self-diagnosis (vs non-autistic) & 0.122 & 1.360 & .176\\
\bottomrule
\end{tabular}
\end{table}

@@

We compared participants who scored above the standard threshold for ``high'' alexithymic traits to those who scored below this threshold on our face processing measures, autistic traits, and anxiety and depression scores. We again removed outliers: this was done using the thresholds used for regression series 1 (see Supplementary Materials for results when all participants including outliers are included). Given that the face processing variables, AQ scores and HADS scores were not normally distributed, these comparisons were conducted using the non-parametric Mann-Whitney U tests. Table \ref{tab:table9} includes the statistics for these comparisons.

\caption{\label{tab:table9}Comparing face processing variables, autistic traits and anxiety and depression symptoms between high and low alexithymia groups.}
\centering
\begin{tabular}[t]{lrrrrl}
\toprule
 & High Alexithymia &  & Low Alexithymia &  & Mann Whitney U, Z,\\
 & Group (Mdn) & N & Group (Mdn) & N & p-value (two-sided)\\
\midrule
Emotion Attribution & 0.1588 & 63 & 0.1411 & 111 & 3082, -1.298, .195\\
Identity Attribution & 0.1 & 63 & 0.0958 & 107 & 3052.5, -1.026, .306\\
Emotion PSE & 0.497 & 63 & 0.4927 & 111 & 3151.5, -1.0803, .281\\
\addlinespace
Identity PSE & 0.5284 & 65 & 0.5287 & 111 & 3348, -0.79548, .427\\
AQ & 88 & 65 & 67 & 113 & 1390.5, -6.897, <.001\\
HADS Anxiety & 9 & 65 & 11 & 113 & 4811.5, 3.4607, <.001\\
HADS Depression & 9 & 65 & 9 & 113 & 3163.5, -1.5557, .12\\
\bottomrule
\end{tabular}
\end{table}

Comparing participants who scored above and below the high alexithymia cut-offs on the face processing variables, no comparisons reached significance (see Table \ref{tab:table9}). Highly alexithymic participants had higher autistic traits, and lower anxiety scores. Depression scores were not significantly different between the group.
